---
title: "2021_w41_nurses"
author: "Albert Rapp"
date: "6 10 2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

Load data and clean names

```{r}
library(tidyverse)
library(paint)
nurses <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-10-05/nurses.csv')

nurses <- nurses %>% janitor::clean_names()
```

Rearrange most of the columns by pivoting

```{r}
nurses <- nurses %>% 
  rename(hourly_50th_percentile = hourly_wage_median) %>% 
  rename(annual_50th_percentile = annual_salary_median) 

percentiles_idx <- nurses %>% 
  colnames() %>% 
  str_detect("percentile") %>% 
  which()


others_idx <- nurses %>% 
  colnames() %>% 
  str_detect("percentile") %>% 
  {!.} %>% 
  which()

percentile_wages <- nurses %>% select(percentiles_idx, state, year) %>% 
  pivot_longer(
    cols = -c("state", "year"),
    names_pattern = "(.+)_(.{2})th_percentile",
    names_to = c("time_frame", "percentile"),
    values_to = "wage"
  ) %>% 
  mutate(percentile = as.numeric(percentile))

cleaned_percentiles <- nurses %>% 
  select(others_idx) %>% 
  left_join(percentile_wages, by = c("state", "year"))
```

Get additional data from [Wikipedia](https://en.wikipedia.org/wiki/List_of_regions_of_the_United_States) to compare regions. 
Text was copied into a csv-file `2021_w41_region_states.csv`.

```{r}
splitting_regex <- "(,\\s){1}(and\\s)?"
splitting_regex2 <- "((\\.|;)\\s){1}(and\\s)?"
split_states <- function(str) {
  if (str_detect(str, "D.C.")) {
    split_regex <- splitting_regex2
  } else {
    split_regex <- splitting_regex
  }
  str %>% str_split(split_regex) %>% unlist()
}

# Copied text from Wikipedia into a csv file and split it using regexes
regions <- read_csv2("2021_w41_region_states.csv", col_names = F) %>% 
  rename(region = X1, state = X2) %>% 
  mutate(state = map(state, split_states)) %>% 
  unnest(state) %>% 
  mutate(state = if_else(state == "Washington, D.C", "District of Columbia", state))

state_region_hourlies <- cleaned_percentiles %>%
  filter(time_frame == "hourly") %>% 
  select(state, year, wage, percentile) %>% 
  left_join(regions, by = "state") %>% 
  mutate(region = if_else(is.na(region), "Other", region))

```


Let's try to use `gt` to make a table.

```{r}
library(gt)
library(gtExtras)
tbl <- state_region_hourlies %>% 
  filter(percentile == 50, year == 2020) %>% 
  arrange(desc(wage)) %>% 
  select(state, wage, region) 

tbl %>% 
  gt(groupname_col = "region") %>% 
  cols_label(
    state = "State", 
    wage = html("Median hourly wage<br>(in USD)")
  ) %>% 
  gt_hulk_col_numeric(wage)
```

First attempt looks interesting.
Let's try to add another column with confidence intervals since we have some quantiles in the data set.
First, create the plots and save them as PNGs.
Make sure they are large enough for later import into the table.

```{r}
hourlies_2020 <- state_region_hourlies %>% 
  filter(year == 2020) %>% 
  mutate(
    new_col_names = case_when(
      percentile == 50 ~ "median",
      percentile == 10 ~ "lower_80",
      percentile == 90 ~ "upper_80",
      percentile == 25 ~ "lower_50",
      percentile == 75 ~ "upper_50",
    )
  ) %>% 
  pivot_wider(
    id_cols = c("state", "region"),
    names_from = new_col_names,
    values_from = wage
  ) 

interval_color1 <- "Dodgerblue1"
interval_color2 <- "Dodgerblue3"
median_color <- "Firebrick4"
range_wages <- state_region_hourlies %>% 
  filter(year == 2020) %>% 
  pull(wage) %>% 
  range()


create_confidence_bar_rect <- function(state) {
  plt <- hourlies_2020 %>% 
    filter(state == !!state) %>% 
    ggplot() +
    geom_rect(
      ymin = 0, ymax = 1,
      aes(xmin = lower_80, xmax = upper_80), 
      alpha = 0.5, 
      fill = interval_color1
    ) +
    geom_rect(
      ymin = 0, ymax = 1,
      aes(xmin = lower_50, xmax = upper_50), 
      size = line_size, 
      alpha = 0.75, 
      fill = interval_color2
    ) +
    geom_rect(
      ymin = 0, ymax = 1,
      aes(xmin = median*.99, xmax = median*1.01), 
      fill = median_color
    ) +
    coord_cartesian(xlim = range_wages, ylim = c(0, 1), expand = T) +
    theme_void()
  
  ggsave(
    glue::glue("2021_w41_nurses_pngs/{state}.png"), 
    plt,
    width = 120,
    height = 5,
    units = "cm"
  )
}

tbl$state %>% walk(create_confidence_bar_rect)
```

Use the pngs and add them to table via `gt_img_rows`.
Add some custom options et voilà.

```{r}
html_text <- '<p style="color:#1874cd; display:inline"; >50%</p> / <p style="color:#1e90ff; display:inline";>90%</p> Confidence-Intervalls and <p style="color:	#8b1a1a; display:inline";>Median</p>'

tbl %>% 
  mutate(state_plot = glue::glue("2021_w41_nurses_pngs/{state}.png")) %>% 
  gt(groupname_col = "region") %>% 
  gt_img_rows(columns = "state_plot", img_source = "local", height = 15) %>% 
  gt_theme_nytimes() %>% 
  cols_label(
    state = "State", 
    wage = html("Median hourly wage<br>(in USD)"),
    state_plot = html(html_text)
  ) %>% 
  gt_hulk_col_numeric(wage) %>% 
  tab_options(
    data_row.padding = px(1),
    row_group.background.color = "grey70",
    row_group.font.weight = "bold",
    column_labels.font.weight = "bold",
    column_labels.font.size = 20
  ) %>% 
  tab_style(
    locations = cells_body(columns = "wage"),
    style = list(
      cell_text(font = "bold", align = "center")
    )
  )
  
```



