---
output: html_document
editor_options: 
  chunk_output_type: console
---

# Packages and Initial Data

```{r}
library(tidyverse)
library(ggbump)
library(patchwork)

camcorder::gg_record(
  dir = here::here('2022_w28_euroflights/img'),
  width = 20,
  height = 20,
  units = 'cm',
  dpi = 600
)
flights <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-07-12/flights.csv') %>% 
  janitor::clean_names()

```

# Colors

```{r}
color_palette <- viridisLite::mako(8)
description_color <- 'grey40'
country_colors <- c(
  'United Kingdom' = color_palette[2],
  'Germany' = color_palette[3],
   'Spain' = color_palette[4],
   'France' = color_palette[5],
   'Italy' = color_palette[6], 
  'Türkiye' = color_palette[7]
)
```

# Basic Bump plot

## Compute data

```{r}
country_flights_by_year <- flights %>% 
  select(year, state = state_name, flights = flt_tot_1) %>% 
  group_by(year, state) %>% 
  summarise(flights = sum(flights))


country_rank_by_year <- country_flights_by_year %>% 
  group_by(year) %>% 
  mutate(
    rank = row_number(desc(flights))
  ) %>% 
  ungroup() %>% 
  arrange(rank, year) 

max_rank <- 6
todays_top <- country_rank_by_year %>% 
  filter(year == 2022, rank <= max_rank) %>% 
  pull(state)
```

## Build bump plot

```{r}
bump_chart_basic <- country_rank_by_year %>% 
  filter(state %in% todays_top) %>% 
  ggplot(aes(year, rank, col = state)) + 
  geom_point(shape = '|', stroke = 6) +
  geom_bump(size = 1) +
  geom_text(
    data = country_rank_by_year %>% filter(year == 2016, state %in% todays_top),
    aes(label = state),
    hjust = 1,
    nudge_x = -0.1,
    fontface = 'bold'
  ) +
  geom_text(
    data = country_rank_by_year %>% filter(year == 2022, state %in% todays_top),
    aes(label = rank),
    hjust = 0,
    nudge_x = 0.1,
    size = 5,
    fontface = 'bold'
  ) +
  annotate(
    'text',
    x = c(2016, 2022),
    y = c(0.25, 0.25),
    label = c(2016, 2022),
    hjust = c(0, 1),
    vjust = 1,
    size = 4,
    fontface = 'bold',
    color = description_color
  ) +
  scale_y_reverse(position = 'right', breaks = seq(16, 2, -2)) +
  scale_color_manual(values = country_colors) +
  coord_cartesian(xlim = c(2014, 2022.5), ylim = c(6.5, 0.25), expand = F) +
  theme_void() +
  theme(
    legend.position = 'none',
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    plot.background = element_rect(fill = color_palette[8], color = NA),
    text = element_text(
      color = description_color,
      family = 'Merriweather'
    ),
    plot.subtitle = element_text(
      margin = margin(t = 3, b = 2, unit = 'mm')
    ),
    plot.title = element_text(
      face = 'bold',
      size = 20
    )
  ) +
  labs(
    title = 'European flight ranking',
    subtitle = 'Rank by number of incoming and outgoing flights from 2016 to 2022'
  )
```

# Ribbon bump plot

## Compute data for bars

Here I have used a little bit of functional programming to compute the percentages for each year.
This means that we have to compute lower and upper bounds for the bars (with `cumsum()`)

```{r}
bar_width <- 0.1
others_color <- 'grey70' # Necessary to not leave out contribution of other countries

compute_percentage_bounds <- function(year) {
  country_rank_by_year %>% 
    filter(year == !!year) %>% 
    arrange(desc(flights)) %>% 
    mutate(
      percentage_flights = flights / sum(flights),
      state = if_else(state %in% todays_top, state, 'Others')
    ) %>% 
    group_by(year, state) %>% 
    summarise(
      percentage_flights = sum(percentage_flights), 
      rank = min(rank),
      .groups = 'drop'
    ) %>% 
    arrange(desc(rank)) %>% 
    mutate(
      percentage_flights_lower = cumsum(c(0, percentage_flights[-7])),
      percentage_flights_upper = cumsum(percentage_flights)
    )
}

percentages_countries_by_year <- map_dfr(2016:2022, compute_percentage_bounds)
```

## Plot Bars

```{r}
percentage_bars_plot <- percentages_countries_by_year %>% 
  ggplot() +
  geom_rect(
    aes(
      xmin = year - bar_width, 
      xmax = year + bar_width,
      ymin = percentage_flights_lower,
      ymax = percentage_flights_upper,
      fill = state
    ),
    col = 'white'
  )+
  theme_void() +
  theme(
    legend.position = 'none',
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    plot.background = element_rect(fill = color_palette[8], color = NA),
    text = element_text(
      color = description_color,
      family = 'Merriweather'
    ),
    plot.subtitle = element_text(
      margin = margin(t = 3, unit = 'mm'),
      lineheight = 1.25
    ),
    plot.title = element_text(
      face = 'bold',
      size = 20
    )
  ) +
  labs(
    title = 'European flight ranking',
    subtitle = 'Rank by number of incoming and outgoing flights from 2016 to 2022.\nHeight of bars indicate percentage of total flights.'
  ) +
  scale_fill_manual(values = c(country_colors, 'Others' = others_color)) 

percentage_bars_plot
```

## Compute Ribbons data

```{r}
n_points <- 100

margin_between_ribbons <- 0.003
compute_ribbons <- function(state) {
  
  state_data <- percentages_countries_by_year %>% 
    filter(state == !!state)
  
  lower_bounds <- state_data %>% 
    select(year, percentage_flights_lower) %>% 
    mutate(
      x_from = year + bar_width, 
      x_to = year + 1 - bar_width,
      y_from = percentage_flights_lower + margin_between_ribbons,
      y_to = c(percentage_flights_lower[-1], percentage_flights_lower[7]) + margin_between_ribbons,
      sigmoid = pmap(list(x_from, x_to, y_from, y_to), sigmoid, n = n_points, smooth = 8)
    ) %>% 
    filter(year < 2022) %>% 
    unnest(sigmoid) %>% 
    select(x, lower = y)
  
  upper_bounds <- state_data %>% 
    select(year, percentage_flights_upper) %>% 
    mutate(
      x_from = year + bar_width, 
      x_to = year + 1 - bar_width,
      y_from = percentage_flights_upper - margin_between_ribbons,
      y_to = c(percentage_flights_upper[-1], percentage_flights_upper[7]) - margin_between_ribbons,
      sigmoid = pmap(list(x_from, x_to, y_from, y_to), sigmoid, n = n_points)
    ) %>% 
    filter(year < 2022) %>% 
    unnest(sigmoid) %>% 
    select(x, upper = y)
  
  left_join(lower_bounds, upper_bounds, by = 'x') %>% 
    mutate(state = state)
}

ribbons_data <- map_dfr(todays_top, compute_ribbons)
```

## Compute mean percentage for labels of rank and country

```{r}
mean_percentages <- percentages_countries_by_year %>% 
  group_by(state, year) %>% 
  summarise(
    middle_percentage = mean(c(percentage_flights_lower, percentage_flights_upper)),
    rank = max(rank),
    .groups = 'drop'
  )
```

## Add ribbons to previous plot

```{r}
combined_plot <- percentage_bars_plot +
  geom_ribbon(
    data = ribbons_data,
    aes(x = x, ymin = lower, ymax = upper, fill = state),
    alpha = 0.5
  ) +
  annotate(
    'text',
    x = c(2016 - bar_width, 2022 + bar_width),
    y = c(1.1, 1.1),
    label = c(2016, 2022),
    hjust = c(0, 1),
    vjust = 1,
    size = 4,
    fontface = 'bold',
    color = description_color
  ) +
  geom_text(
    data = mean_percentages %>% filter(year == 2016),
    aes(x = year - bar_width, y = middle_percentage, label = state, col = state),
    hjust = 1,
    nudge_x = -0.1,
    fontface = 'bold'
  ) +
  geom_text(
    data = mean_percentages %>% filter(year == 2022, state %in% todays_top),
    aes(x = year + bar_width, y = middle_percentage, label = rank, col = state),
    hjust = 0,
    nudge_x = 0.1,
    size = 5,
    fontface = 'bold'
  ) +
  coord_cartesian(
    xlim = c(2014.5, 2022.25),
    ylim = c(0, 1.1)
  ) +
  scale_color_manual(values = c(country_colors, "Others" = others_color)) 

combined_plot
```

## Turn order of layers so that bars are always on top

```{r}
ribbons <- combined_plot$layers[[2]] 
rects <- combined_plot$layers[[1]]

combined_plot$layers[[1]] <- ribbons 
combined_plot$layers[[2]] <- rects
ribbon_bump_chart <- combined_plot
```


# Combine both bumpcharts into one plot

```{r}
bump_chart_basic /
  plot_spacer() /
  ribbon_bump_chart +
  plot_layout(heights = c(0.45, 0.02, 0.49)) +
  plot_annotation(
    title = 'Basic bumps vs. ribbon bump charts',
    caption = 'TidyTuesday 2022 - Week 28 | Data: Eurocontrol | Graphic: Albert Rapp, @rappa753',
    theme = theme(
      plot.title = element_text(
        size = 32,
        color = description_color,
        family = 'Oleo Script',
        face = 'bold'
      ),
      plot.caption = element_text(
        size = 8,
        color = description_color,
        family = 'Oleo Script',
        face = 'bold'
      )
    )
  ) &
  theme(
     plot.background = element_rect(fill = color_palette[8], color = NA)
  )
```

